Dynamic Stochastic Approximation Monte Carlo

### 2018-06-25 papers/histogram/sad.tex
spell checks in paper

Around the same time, the convergence rate for WL was being
explored~\cite{zhou2005understanding,lee2006convergence,
belardinelli2007wang}. It was found that utilizing modification factors
that decrease faster than $1/t^2$ leads to
nonconvergence~\cite{belardinelli2007fast}.  This lead to the so-called
$1/t$-WL algorithm which ensures that excess CPU time is not wasted by
continuing to perfrom calculations once the error in the density of
states becomes saturated~\cite{belardinelli2008analysis}. An effort to
reduce the number of user defined parameters of WL and formulate a
theory for why the method converges despite detailed balance being
violated although infrequently was also being udertaken.

@@

The Wang-Landau Algorithm developed by Wang and Landau proposes a
random walk in energy space in order to estimate the density of
states~\cite{wang2001efficient,wang2001determining, landau2014guide}.
Wang-Landau's major tenant is that when counting the histogram
$H(\epsilon)$, the energy occurences should form a flat distribution.
The criteria for flat sampling is given by
\begin{equation}
	\frac{\min_{\epsilon} H(\epsilon)}
	{\big\langle H(\epsilon)\big\rangle }
	> \eta
\end{equation}
where $\eta$ is usually set to $0.80$ and determines how many times
each energy is sampled relative to the mean number of visits.
Difficulties can occur with this flatness criteria due to the fact that
some energies on a given energy range might never be
sampled~\cite{haber2014transition}. The algorithm is briefly outlined
below where $w$ represents the weights:
\begin{equation}
	\ln{D_{i+1}(\epsilon,N)}=\ln{D_{i}(\epsilon,N)}
	+\gamma
\end{equation}
\begin{equation}
	\gamma_{i+1}=\frac{u}{2} \gamma_{i}
\end{equation}
for $u$ typically greater than 1 and $\gamma$ is the modification
factor for the density of states.  A new configuration for the density
of states with a probability given by the following is chosen:
\begin{equation}
	\mathcal{P}(\epsilon_\text{old} \rightarrow \epsilon_\text{new})
	= \min[1,e^{\ln{D(\epsilon_\text{old})}-\ln{D(\epsilon_\text{new})}}]
\end{equation}
The entire process is repeated until a desired cutoff is reached.

@@

\section{SAD Algorithm}
\subsubsection{The theory behind SAD}
The SAD (Dynamic Stochastic Approximation) is a version of the SAMC
Algorithm that attempts to dynamically choose the modification factor
rather than relying on system dependent parameters such as $t_0$ or
$\gamma_0$.  There is an immediate advantage of such an algorithm where
parameters are chosen independent of system size or type. Each
Broad-Histogram method has unique advantages and disadvantages.
Wang-Landau requires an energy range for initialization.  SAMC removes
this energy range requirement but requires simulating every possible
energy. Our proposed method SAD requires the user to input
$T_\text{min}$ which is an immediate disadvantage of the method;
however, this allows the simulating to define an energy simulation
range between $T_\text{min}$ and $\infty$. Setting a physical parameter
such as a minimum temperature $T_\text{min}$ is intuitive vs. a user
needing to know a prior either an energy range or some unphysical
parameter $t_0$.

>>>>

Around the same time, the convergence rate for WL was being
explored~\cite{zhou2005understanding,lee2006convergence,
belardinelli2007wang}. It was found that utilizing modification factors
that decrease faster than $1/t^2$ leads to
nonconvergence~\cite{belardinelli2007fast}.  This leads to the so-called
$1/t$-WL algorithm which ensures that excess CPU time is not wasted by
continuing to perform calculations once the error in the density of
states becomes saturated~\cite{belardinelli2008analysis}. An effort to
reduce the number of user defined parameters of WL and formulate a
theory for why the method converges despite detailed balance being
violated although infrequently was also being undertaken.

@@

The Wang-Landau Algorithm developed by Wang and Landau proposes a
random walk in energy space in order to estimate the density of
states~\cite{wang2001efficient,wang2001determining, landau2014guide}.
Wang-Landau's major tenant is that when counting the histogram
$H(\epsilon)$, the energy occurrences should form a flat distribution.
The criteria for flat sampling is given by
\begin{equation}
	\frac{\min_{\epsilon} H(\epsilon)}
	{\big\langle H(\epsilon)\big\rangle }
	> \eta
\end{equation}
where $\eta$ is usually set to $0.80$ and determines how many times
each energy is sampled relative to the mean number of visits.
Difficulties can occur with this flatness criteria due to the fact that
some energies on a given energy range might never be
sampled~\cite{haber2014transition}. The algorithm is briefly outlined
below where $w$ represents the weights:
\begin{equation}
	\ln{D_{i+1}(\epsilon,N)}=\ln{D_{i}(\epsilon,N)}
	+\gamma
\end{equation}
\begin{equation}
	\gamma_{i+1}=\frac{u}{2} \gamma_{i}
\end{equation}
for $u$ typically greater than 1 and $\gamma$ is the modification
factor for the density of states.  A new configuration for the density
of states with a probability given by the following is chosen:
\begin{equation}
	\mathcal{P}(\epsilon_\text{old} \rightarrow \epsilon_\text{new})
	= \min[1,e^{\ln{D(\epsilon_\text{old})}-\ln{D(\epsilon_\text{new})}}]
\end{equation}
The entire process is repeated until a desired cutoff is reached.

@@

\section{SAD Algorithm}
\subsubsection{The theory behind SAD}
The SAD (Dynamic Stochastic Approximation) method is a version of the SAMC
Algorithm that attempts to dynamically choose the modification factor
rather than relying on system dependent parameters such as $t_0$ or
$\gamma_0$.  There is an immediate advantage of such an algorithm where
parameters are chosen independent of system size or type. Each
Broad-Histogram method has unique advantages and disadvantages.
Wang-Landau requires an energy range for initialization.  SAMC removes
this energy range requirement but requires simulating every possible
energy. Our proposed method SAD requires the user to input
$T_\text{min}$ which is an immediate disadvantage of the method;
however, this allows the user to define an energy simulation
range between $T_\text{min}$ and $\infty$. Setting a physical parameter
such as a minimum temperature $T_\text{min}$ is intuitive vs. a user
needing to know a prior either an energy range or some unphysical
parameter $t_0$.
